Argennon: A Scalable Smart Contract Platform

### 2022-03-29 tex/sections/avm.tex
Save the draft of Ch1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

An Argennon Standard Representation (ASR) is a programming language for describing argennon applications, optimized
for the architecture and requirements of the Argennon platform.
Argennon supports two standard representations, one is a high level text based language which needs
costly compilation before being executed on a hardware machine. The other one is a low level binary representation which
usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
language is intended for preserving high level information of applications logic to facilitate
platform specific compiler optimization at host nodes. On the other hand, the low level language enables efficient
execution of applications that are not frequently used.

@@

The Argennon Smart Contract Execution Environment can be seen as a machine for executing http requests, producing
their http response and updating related heap
chunks. The AscEE executes requests sequentially \footnote{Actually requests are executed in parallel but by
performing data dependency analysis the result is guaranteed to be identical with sequential execution of
requests.} and each request has a separate \emph{execution session}. Execution sessions are separate sessions of
executing smart contract's code by the Argennon Smart Contract Execution Environment. The state of an execution
session will be
destroyed at the end of the session and only the state of heap chunks is preserved. If a session fails and does not
complete normally, it will not have any effect on the state of heap chunks.

During an execution session an application can make http requests to other applications. Those requests will not start
a new execution session and will be executed within the current session. In AscEE making a http request is similar to
a function invocation, because of that, we sometimes refer to it as an application
call.

@@

Argennon has three primitive prefix trees:
\emph{applications, accounts} and \emph{local}.
All these trees are in base 256, with the maximum height
of 8. An Argennon identifier may be simple or compound. A simple identifier is generated using a single trie, while a
compound identifier is generated by concatenating prefix codes generated by two or more tries:

@@

\begin{itemize}
    \item \texttt{read\_only}: only allows read and check operations.
    \item \texttt{writable} memory locations inside this access block can not be modified.
    \item \texttt{check\_only}: only allows check operations. These operations query the persistence
    status of memory locations.
    \item \texttt{additive}: only allows addition operations without overflow checking. Note that the content of these
    access blocks cannot be read.
\end{itemize}

Chunks are intended for simplifying proof checking of the data stored in the Argennon cloud and access blocks are
required for better parallelization of request execution. An application should put the data it predicts is needed for
validating a block in the same chunk and put the data it predicts will be needed in a single execution session in
the same access block.

@@

Completing an execution session needs computational resources. The amount of resources used by an execution session
needs to be monitored, otherwise a spammer would be able to easily exhaust the resources of the execution
environment. Resource usage can be measured per session or per application call.

The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
session is always measured per session and default pre-defined resource caps are used. On the other hand in a monitored
session some resources are measured per application and some
caps are determined dynamically by the request. An optimistic session must always
complete normally but a monitored session is allowed to either complete normally or abruptly.

Obviously resource monitoring is easier for optimistic sessions, and they can be executed more efficiently. However, if
an optimistic session violates its resource caps, determining the point of failure requires precise resource
measurement. Note that for implementing optional decoupling principle, we need to determine the exact application
which has failed in the call chain. For example assume that an optimistic session containing an application call
violates a 2 milliseconds execution time cap. If in the caller application the call happens exactly after 2
milliseconds of execution, a small fluctuation in the execution time measurement can
change the point of failure between the called and the main application. This will introduce nondeterministic
behaviour which makes block validation impossible. On the other hand, for a monitored session this is not an issue
because resource measurements and caps are defined per application, and we make sure that resource caps are not too
small for any application call.

@@

    Optimistic sessions have a predefined \texttt{maxClocks} value which is determined by the Argennon protocol. This
    value defines a bound on the total cpu time of the session and no per application measurement is done.

    Monitored sessions perform per application cpu-time measurement, and every application call during a monitored
    session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
    can be used for executing that particular application call. It should be noted that the timer is paused when an
    application makes a call to another application, until the control returns to the caller. An application call
    needs at least 100 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
    considered a failed call.

@@

    The amount of external clocks for every application
    call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result the total number of clocks of a monitored
    session is always less than \(3 \times \texttt{maxClocks}\) of the root application. The value of
    \texttt{maxClocks} for
    the root application call is determined by the transaction.

    \item \textbf{local memory}:
    any memory usage of an application that is not part of the heap chunks and is not part of another resource
    will be considered as local memory usage. This resource is measured in bytes.
    \item \textbf{heap access list}:
    every session can only access heap locations that are declared in its access list. In addition,
    resizing heap chunks can only be done in the range of the pre-declared lower bound and upper bound.
    \item \textbf{applications list}:
    a session may only make requests to applications that are declared in its application list.
    \item \textbf{call depth}:
    the number of nested application calls can not be more than a threshold. This threshold is determined by the
    Argennon protocol.
    \item active differed calls:
    \item virtual signatures:
    \item number of locks

@@

>>>>

An Argennon Standard Representation (ASR) is a programming language for describing argennon applications, optimized
for the architecture and requirements of the Argennon platform.
Argennon supports two standard representations: one is a high level text based language which needs
costly compilation before being executed on a hardware machine. The other one is a low level binary representation which
usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
language is intended for preserving high level information of applications logic to facilitate
platform specific compiler optimization at host nodes. On the other hand, the low level language enables efficient
execution of applications that are not frequently used.

@@

The Argennon Smart Contract Execution Environment can be seen as a machine for executing http requests, producing
their http response and updating related heap
chunks. The AscEE executes requests sequentially \footnote{Actually requests are executed in parallel but by
performing data dependency analysis the result is guaranteed to be identical with sequential execution of
requests.} and each request has a separate \emph{execution session}. Execution sessions are separate sessions of
executing smart contract's code by the Argennon Smart Contract Execution Environment. The state of an execution
session will be
destroyed at the end of the session and only the state of heap chunks is preserved. If a session fails and does not
complete normally, it will not have any effect on the heap chunks.

During an execution session an application can make http requests to other applications. Those requests will not start
a new execution session and will be executed within the current session. In AscEE making a http request to an
application is similar to a function invocation therefore we also refer to it as an application call.

@@

Argennon has four primitive prefix trees:
\emph{applications, accounts, local} and \emph{varUint}.
All these trees are in base 256, with the maximum height
of 8. An Argennon identifier may be simple or compound. A simple identifier is generated using a single trie, while a
compound identifier is generated by concatenating prefix codes generated by two or more tries:

@@

\begin{itemize}
    \item \texttt{read\_only}: only allows read and check operations.
    \item \texttt{writable}: memory locations inside this access block can not be modified.
    \item \texttt{check\_only}: only allows check operations. These operations query the persistence
    status of memory locations.
    \item \texttt{additive}: only allows addition operations without overflow checking. Note that the content of these
    access blocks cannot be read.
\end{itemize}

Chunks are intended for simplifying proof checking of the data stored in the Argennon cloud and access blocks are
required for better parallelization of the request execution. An application should put the data it predicts is
needed for validating a block in the same chunk and the data it predicts is needed in a single execution
session in the same access block.

@@

Completing an execution session needs computational resources. The amount of resources used by an execution session
should be monitored, otherwise a malicious user would be able to easily spam and exhaust resources of the execution
environment. Resource usage can be measured per session or per application call.

The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
session is always measured per session and default pre-defined resource caps are used. On the other hand in a monitored
session some resources are measured per application and some
caps are determined dynamically by the transaction. An optimistic session must always
complete normally, a monitored session is allowed to either complete normally or abruptly.

Obviously resource monitoring is easier and more efficient for optimistic sessions. However, if
an optimistic session violates its resource caps, determining the point of failure requires precise resource
measurement. Note that for implementing optional decoupling principle, we need to determine the exact application
which has failed in the call chain. For example assume that an optimistic session containing an application call
violates a 2 milliseconds execution time cap. If in the caller application the call happens exactly after 2
milliseconds of execution, a small fluctuation in the execution time measurement can
change the point of failure between the called and the main application. This can introduce nondeterministic
behaviour which makes block validation impossible. On the other hand, for a monitored session this is not an issue
because resource measurements and caps are defined per application, and we make sure that resource caps are not too
small and can be measured easily.

@@

    Optimistic sessions have a predefined \texttt{maxClocks} value which is determined by the Argennon protocol. This
    value defines a bound on the \textbf{total} cpu time of the session and no per application measurement is done.

    Monitored sessions perform per application cpu-time measurement, and every application call during a monitored
    session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
    can be used for executing that particular application call. It should be noted that this cpu timer is paused when an
    application makes a call to another application, until the control returns to the caller. An application call
    needs at least 100 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
    considered a failed call.

@@

    The amount of external clocks of an application
    call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result, the total number of clocks of a monitored
    session is always less than \(3 \times \texttt{maxClocks}\) of the root application. The value of
    \texttt{maxClocks} for
    the root application call is determined by the transaction.

    \item \textbf{local memory}:
    any memory usage of an application that is not part of a heap chunk and is not part of another resource
    will be considered as local memory usage. Local memory is not persistent and when an application finishes serving a
    request and returns the http response (i.e.\ the application call completes) its local memory is deleted.
    This resource is measured in bytes.

    Optimistic sessions measure local memory usage per session and enforce a protocol-defined cap on the total amount of
    used local memory in the session. Monitored sessions measure local memory usage per
    application call and enforce a protocol defined-cap per application call. An application call which tries to use
    more local memory than the cap will fail.
    \item \textbf{heap access list}:
    every session can only access heap locations that are declared in its access list. In addition,
    resizing heap chunks can only be done in the range of the pre-declared lower bound and upper bound.
    \item \textbf{applications list}:
    a session may only make requests to applications that are declared in its application list.
    \item \textbf{call depth}:
    during a session the number of nested application calls can not be more than a threshold. This threshold is
    determined by the Argennon protocol.
    \item active differed calls:
    \item virtual signatures:
    \item number of locks

@@

The Argennon protocol allows AscEE's implementations to have at max \%50 error in measuring execution time and at
max \%25 error in measuring local memory usage. Other resources must be measured precisely.

\section{Request Attachments}\label{sec:attachments}

Attachments of a request is a list of request identifiers of the current block that are "attached" to the request.
That means, for validating that request a validator first needs to "inject" the digest of attached requests into the
httpRequest field of the request.

The main usage of this feature is for fee payment. A request that wants to pay the fees for a number of requests
declares those requests as its attachments. For paying fees the payer signs the digest of requests for which he
wants to pay fees. After injecting the digest of those request by validators, that signature can be validated
correctly and securely.
